---
title: "SQL"
output: html_document
---

## Filtering Rows and Columns

In SQl you can filter columns by using SELECT and rows by using WHERE. Let's look at an example using the Lahman baseball database. First, we should load the Lahman package. We will also need the sqldf package.

```{r message=FALSE, warning=FALSE}
library(Lahman)
library(sqldf)
```

Let's take a look at an example. Suppose we want to see the homerun totals for the 1927 Yankees. We could write: 

```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting
WHERE teamID='NYA' and yearID=1927"
sqldf(query)
```

## ORDER BY

With SQL, you can also order the rows in the output. For example, suppose you wanted to see every instance of a player hitting more than 50 homeruns, but you would like players with the most homeruns to be at the top of this list. We could do this in the following way:

```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting
WHERE HR>50
ORDER BY HR DESC"

sqldf(query)
```

Let's look at another example of ways to order your data. 

 

### Example 

 

Question 1: Find all instances of a player striking out less than 10 times. Make sure each player has at least 400 at bats. Order strikeouts from least to greatest. 

 

```{r} 

query<-"SELECT playerID,yearID,teamID,SO,AB FROM Batting 

WHERE SO<10 and AB>400 

ORDER BY SO" 

sqldf(query) 

``` 

 

## Aggregation 

 

Aggregatiion allows one to group and condense information in different ways, essentially 'squashing' different pieces of data together. This is done by using the GROUP BY command. When using GROUP BY, one must then specify how to see the minimum or maximum. These modifiers are placed on the other column names. This is how one would specify what should be done to those data sets:  

*Summation: sum(colname) 

*Average: avg(colname) 

*Maximum: max(colname) 

*Minimum: min(colname) 

If one wants to continue filtering after they have grouped, they can use the HAVING command, which functions as the filtering WHERE command when used after GROUP BY. ORDER BY can be used in the same way as before, coming at the end of the command list. 

 

### Examples 

Question 1: Find the career homerun totals for Babe Ruth. 

 

```{r} 

query<-"SELECT playerID,sum(HR) FROM Batting 

WHERE playerID='ruthba01' 

GROUP BY playerID" 

sqldf(query) 

``` 

 

Question 2: Find the career totals for all players but list only those with a total of 600 or more. Order by having greatest total at the top. 

 

```{r} 

query<-"SELECT playerID,sum(HR) FROM Batting 

GROUP BY playerID 

HAVING sum(HR)>=600 

ORDER BY sum(HR) DESC" 

sqldf(query) 

``` 

 

Question 3: What players have the highest average per season homerun totals? List only averages greater than 30. Order from greatest to least. 

 

```{r} 

query<-"SELECT playerID,avg(HR) FROM Batting 

GROUP BY playerID 

HAVING avg(HR)>30 

ORDER BY avg(HR) DESC" 

sqldf(query) 

``` 

 

## Joining 

 

Joining allows tables to be joined together based on common data to increase the data output options. This is done through the INNER JOIN and ON commands. INNER JOIN indicates which tables are being joined together. ON indicates the common data that the tables will be joined by. If the common data appears alone in the code, it will need to either have specified which of the joined tables it is coming from, or be replaced by the information from another table. In the first example, playerID in the SELECT phrase was replaced by nameFIRST and nameLAST to make reading the table easier. 

 

### Examples 

Question 1: First and last names along with teamID, yearID, and homeruns for Babe Ruth. 

 

```{r} 

query<-"SELECT nameFIRST,nameLAST,teamID,yearID,HR  
FROM Batting INNER JOIN Master 
ON Batting.playerID=Master.playerID 
WHERE Batting.playerID='ruthba01'" 

sqldf(query) 

``` 

 

Question 2: Find all instances of players hitting more than 50 homeruns. Include first and last names, team, year, and homeruns in descending order. 

 

```{r} 

query<-"SELECT nameFirst,nameLast,teamID,yearID,HR 
FROM Batting INNER JOIN Master 
ON Batting.playerID=Master.playerID 
WHERE Batting.HR>50
ORDER BY HR DESC" 

sqldf(query) 
``` 

 

Question 3: Find Babe Ruth's homeruns by season, with the team name displayed. 

 

```{r} 

query<-"SELECT playerID,name,Batting.yearID,Batting.HR 

FROM Batting INNER JOIN Teams 

ON Batting.teamID=Teams.teamID and Batting.yearID=Teams.yearID 

WHERE playerID='ruthba01'" 

sqldf(query) 

``` 